---
title: "NHDD"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

Indeed, data analyzed by The Washington Post shows that the housing recovery has been strongest in the wealthiest areas, and slowest in the poorest. The average price of a house in a zip code in the top 10 percent of the wealth distribution rose more than 20 percent between 2004 and 2015, versus just 13 percent in the rest.

Home values at the high and low ends of the market are now further apart than they’ve ever seen over this time.

## Group by neighborhood stuff

```{r}
# library for sample_n
library('dplyr')
library('leaps')

folder <- 'geos'
files <- dir(folder, pattern=".csv")
alldf <- list()
for (i in 1:length(files)) {
  student <- gsub("([^_]*)_.*.csv", "\\1", files[i])
  alldf[[student]] <- read.csv(paste0(folder, "/", files[i]), as.is=TRUE)
}

dd <- read.csv('final_nhh.csv', as.is = TRUE)
d <- sample_n(dd, 5000)

# turn each sale into its own observation
df <- data.frame(matrix(nrow = 1, ncol = 16))
insert.cols <- c('pid', 'yearbuilt', 'totval', 'bedrooms', 'bathrooms', 'halfbaths', 'landval', 'zone', 'acres', 'exval', 'pctgood', 'sqft', 'garagesqft', "saleprice", "saledate", "saleowner")
colnames(df) <- insert.cols

for (i in seq(1:nrow(d)) ) {
  # for each sale in row
  for (sale in 1:5) {
    # create new row with the right data
    access.cols <- c('pid', 'yearbuilt', 'totval', 'bedrooms', 'bathrooms', 'halfbaths', 'landval', 'zone', 'acres', 'exval', 'pctgood', 'sqft', 'garagesqft', paste(c("saleprice", "saledate", "saleowner"), sale, sep = ""))
    
    r_ <- d[i, access.cols]
    names(r_)
    
    colnames(r_) <- insert.cols
    df <- rbind(df, r_)
    
  }
}

# turn the dates into separate cols via lubridate
library('lubridate')
lbd_helper1 <- mdy(df$saledate)
df$salemonth <- as.factor(month(lbd_helper1))
df$saleyear <- year(lbd_helper1)

```

Data cleaning:
```{r}
# drop the na sales ones
df.na <- df[-which(is.na(df$saledate)),]

df.na$saleprice <- as.numeric(df.na$saleprice)

# drop the ones in which the sale price was 0
df.sale <- df.na[-which(df$saleprice == 0), ]
```

Adding the lat and longitude data
```{r}
# add latitude and longitude data
adf <- data.frame(matrix(ncol=3))
colnames(adf) <- c('pid', 'latitude', 'longitude')
for (d in alldf) {
  newdf <- data.frame(d$pid)
  colnames(newdf) <- c('pid')
  newdf$latitude <- ifelse('latitude' %in% colnames(d), d$latitude, d$lat)
  newdf$longitude <- ifelse('longitude' %in% colnames(d), d$longitude, d$long)
  adf <- rbind(adf, newdf)
}

# some people messed up the lat's and long's 
# get all the ones where the latitude is 
df.loc <- merge(adf, df.sale, by = 'pid')
```

separate into res vs com:
```{r}
# separate into commercial vs residential units
df.com <- df.loc[which(df.loc$bedrooms == 0), ]
df.res <- df.loc[which(df.loc$bedrooms != 0), ]

# get rid of all property transfers
df.res.s <- df.res[-which(df.res$saleprice == 0), ]
```

## Analyzing the change in sale prices by neighborhood
```{r}
# get average price for zone by each year

# dataframe to store stuff
avg.nghbd.price <- data.frame(
  matrix(
    ncol = length(unique(df.res.s$saleyear)),
    nrow = length(unique(df.res.s$zone))
    )
  )
colnames(avg.nghbd.price) <- as.character(unique(df.res.s$saleyear))
rownames(avg.nghbd.price) <- unique(df.res.s$zone)

for (zone in unique(df.res.s$zone) ) {
  for (year in unique(df.res.s$saleyear)) {
    m <- mean(df.res.s$saleprice[which(df.res.s$saleyear == year & df.res.s$zone == zone)], na.rm = TRUE)
    avg.nghbd.price[zone, as.character(year)] <- m
  }
  
}
```
Okay, now for some graphs
```{r}
plot(log(avg.nghbd.price['RS2',]))
```

## Counfounding parameters
- Zoning law changes
- Immigration
